---
title: "SHD_LL97_LL84"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Methodology:

Initial steps prior to analyzing and post-processing in R/R Studio were to download and merge the datasets in Excel.
LL84/09 (amended via LL133/16) data for the 2017 reporting year were obtained from the NYC Mayor's Office of Sustainability website, [Green Buildings & Energy Efficiency](https://www1.nyc.gov/html/gbee/html/plan/ll84_scores.shtml).
A .csv file listing all properties in NYU Furman Center's [Subsidized Housing Database](http://coredata.nyc/) was obtained.
INDEX-MATCH for the reference Building-Block-Lot (BBL) number for each property was performed in Excel to merge the 2 datasets.  Not all subsidized housing projects are required to report whole-building annual energy and water consumption under the local legislation.  Covered buildings include only those exceeding 25,000 sq. ft. (or multiple buildings/condos on same tax lot togethering exceeding 100,000 sq. ft.) and three stores in height.

```{r}
library(maps)
library(dplyr)
library(ggplot2)
library(ggmap)
library(gridExtra)

# Set working directory
setwd("~/Desktop/BLOG/Project1")

# Read in data & define column data types
df_og <- read.csv("full_LL84_SHD_analysis.csv", header = TRUE, na.strings = "NA", colClasses = c(rep("character",2),"numeric", rep("character",9), rep("numeric",2), "character", rep("numeric",2), rep("character",2), rep("numeric",2), rep("character",4),rep("numeric",4), rep("character",9), rep("numeric",2),rep("character",9), rep("numeric",18), rep("character",2)))


# Create copy of data frame; preserve initial import.
dfO <- df_og

# Add field for expiry year using date of program expiration.
for (i in 1:nrow(dfO))
{
  dfO$end_year[i] <- substring(dfO$end_date[i],1,4)
}

# Filter for only those entries that have data per LL84/09.  Use conditional expressions on GFA and source EUI attributes.
dfO <- dfO[(dfO$GFA>0 & dfO$source_EUI>0),]
dfComplete <- dfO[complete.cases(dfO[,c("source_EUI")]),]

# EXPLORE INVENTORY
# Summarize # of units by end date
quantity <- dfComplete %>%
  group_by(sub_subsidy_name) %>%
  summarize(tot_sub_units=sum(subsidized_units))
quantity

# Summarize # of properties & assessed value by end date
expire <- dfComplete %>%
  group_by(end_year) %>%
  summarize(N=n(),total_value=sum(assessed_value))
expire

# Summarize by subsidy program type
program <- dfComplete %>%
  group_by(sub_subsidy_name) %>%
  summarize(N=n()) %>%
  mutate(pct=round((N*100/sum(N)),0))
program

# Create map of non-exempt properties and LL97 financial exposure in 2024
p <- ggplot(data = subset(dfComplete, (GFA>0 & pct_aff<0.35 & end_year<2031)), aes(x = longitude, y = latitude))
p + geom_point(aes(color=X2024_exceed)) + coord_map(projection = "albers", lat0=40.5, lat1=50) +
  guides (fill=FALSE) +
  theme_minimal() + theme(axis.text=element_blank(), panel.grid=element_blank(), legend.text=element_text(size=8), legend.title=element_text(size=9), axis.title=element_blank())

# Create map of non-exempt properties and LL97 financial exposure in 2030
p2 <- ggplot(data = subset(dfComplete, (GFA>0 & pct_aff<0.35 & end_year<2031)), aes(x = longitude, y = latitude))
p2 + geom_point(aes(color=X2030_exceed)) + coord_map(projection = "albers", lat0=40.5, lat1=50) +
  guides (fill=FALSE) +
  theme_minimal() + theme(axis.text=element_blank(), panel.grid=element_blank(), legend.text=element_text(size=8), legend.title=element_text(size=9), axis.title=element_blank())


# Create new dataset with subset of attributes to be used for tileset.
clean_df <- subset(dfComplete, select=c("ref_bbl","data_source","data_year","agency_name","project_name",
                                        "sub_subsidy_name","preservation","start_date","end_date","end_year",
                                        "owner_type","total_units","subsidized_units","address",
                                        "pct_aff","year_built","assessed_value","GFA",
                                        "longitude","latitude","source_EUI","site_EUI",
                                        "nat_gal","electricity","CO2_LL97.19","CO2_CAP_2029",
                                        "CO2_CAP_2034"))

#head(clean_df)

# Assign IDs to classify properties.  
# Cat 0 is for properties that have, as of 2018, operational carbon below the LL97/19-set 2024 emissions limit.
# Cat 1 is for exempt properties (>35% subsidized units) exceeding their carbon limit but thus not incurring fines.
# Cat 2 is for non-exempt properties exceeding their carbon limit and thus financially exposed to penalties.
# Cat 3 is for properties with missing data, i.e., expiry year, total unit/subsidized unit count.

clean_df$cat <- c(rep(3,nrow(clean_df)))
clean_df$cat2 <- c(rep(3,nrow(clean_df)))

for (i in 1:nrow(clean_df)) {
  if (!is.na(clean_df$pct_aff[i]) && !is.na(clean_df$end_year[i])) {
    
    clean_df$cat[i] <- 0
    clean_df$cat2[i] <- 0
    
    if (clean_df$CO2_CAP_2029[i] < clean_df$CO2_LL97.19[i]) {
      if (clean_df$pct_aff[i] >= 0.35) {
        clean_df$cat[i] <- 1
      }
      else if (clean_df$pct_aff[i] < 0.35) {
        clean_df$cat[i] <- 2
      }
    }
    if (clean_df$CO2_CAP_2034[i] < clean_df$CO2_LL97.19[i]) {
      if (clean_df$pct_aff[i] >= 0.35) {
        clean_df$cat2[i] <- 1
      }
      else if (clean_df$pct_aff[i] < 0.35) {
        clean_df$cat2[i] <- 2
      }
    }
  }
}

# Summarize by # of properties in each cat (for future next step: by borough)
catSum <- clean_df %>%
  group_by(cat) %>%
  summarize(N=n()) %>%
  mutate(pct=round((N*100/sum(N)),0))
catSum

# Summary, total # of units set to expire within each compliance period.  Grouped by operational carbon performance
catSum2 <- clean_df %>%
  group_by(cat, r=cut(as.numeric(end_year), breaks= seq(2019, 2050, by = 5))) %>%
  summarize_at(vars(subsidized_units),sum)
catSum2

# Summary, total # of properties with expiring affordability restrictions within each compliance period.  Grouped by operational carbon performance
unitSum <- clean_df %>%
  group_by(cat, gr=cut(as.numeric(end_year), breaks= seq(2019, 2050, by = 5))) %>%
  summarize(N=n())
unitSum


# Export .csv for upload to Mapbox.
write.csv(clean_df, file = "/Users/claudiamezey/Desktop/BLOG/SHD_LL84_2017_inv.csv") 

```

```{r}
library(RColorBrewer)


# experiment with plotting directly in R

register_google(key = "AIzaSyBvsEQw75VL4FETLEjZIdRKDuRNpxwnb-I")
# Create base map of NYC
nyc_map <- get_map(location = c(lon = -74.00, lat = 40.71), maptype = "satellite", zoom = 11, color="bw")
nyc_map <- ggmap(nyc_map)
nyc_map_1 <- nyc_map + stat_density2d(data=subset(clean_df,end_year<2031 & cat!=0), aes(x=longitude, y=latitude, fill=..level.., alpha=..level..), geom="polygon")
## Define the spectral colors to fill the density contours
nyc_map_1 <- nyc_map_1 + scale_fill_gradientn(colours=rev(brewer.pal(7, "Spectral")))
nyc_map_2

# try another
# density is points per unit area
nyc_map_2 <- nyc_map + stat_density2d(data=subset(clean_df,end_year<2031 & cat==1), aes(x=longitude, y=latitude, fill=..level.., alpha=..level..), geom="polygon") +
  scale_fill_gradientn(colours=rev(brewer.pal(7, "Spectral")))

nyc_map_2


#nyc_map <- nyc_map + geom_point(data = subset(clean_df,end_year>=2031),
                           #        aes(x=longitude,y=latitude,
                            #           fill="#56B4E9"),size=1.4,shape=21,stroke=0.2,color="black")
#nyc_map + geom_point(data = subset(clean_df,end_year<2031),aes(x=longitude,y=latitude,
        #                               fill=factor(cat)),size=1.4,shape=21,stroke=0.2,color="black") +
 # theme_minimal() + theme(axis.text=element_blank(), panel.grid=element_blank(), legend.text=element_text(size=8), legend.title=element_text(size=9), axis.title=element_blank()) + scale_fill_manual(values=c("#56B4E9","#E69F00","#c92614"))

```


